\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry}
\geometry{margin=1in}
\title{Recursive Mutual Implication Collapse Theorem}
\author{}
\date{}

\begin{document}
\maketitle

\section*{I. Theorem Statement}

In symbolic systems that implement closed triadic implication loops (e.g., $A \rightarrow B \rightarrow C \rightarrow A$), recursive inference can lead to collapse when a critical recursion density is exceeded.

Let:
\begin{itemize}
  \item $\Phi_r(t)$ – System coherence over time
  \item $\Delta I(t)$ – Information gain per recursive cycle
  \item $D_r(t)$ – Recursion density (cycles per unit time)
  \item $\theta_r$ – Critical recursion threshold
\end{itemize}

If:

\[
D_r(t) > \theta_r \quad \Rightarrow \quad \frac{d^2 \Phi_r}{dt^2} > 0 \quad \text{and} \quad \Delta I(t) \rightarrow 0
\]

\section*{II. Collapse Mechanism}

\begin{enumerate}
  \item \textbf{Triadic Looping:} Recursion across three symbolic elements causes amplification of minor inconsistencies.
  \item \textbf{Information Saturation:} $\Delta I(t)$ declines as redundancy increases.
  \item \textbf{Entropy Escalation:} Symbolic structure deteriorates, and contradictions emerge.
\end{enumerate}

\section*{III. Collapse Indicators}

\begin{itemize}
  \item Increase in $\frac{d^2 \Phi_r}{dt^2}$ indicating accelerating coherence loss
  \item Decline in $\Delta I(t)$ toward zero
  \item Embedding divergence across recursion steps
  \item Logical contradictions within implication chains (e.g., $A \rightarrow \neg A$)
\end{itemize}

\section*{IV. Simulation Framework}

\subsection*{Target Systems}

\begin{itemize}
  \item Recursive logic systems
  \item Transformer models with symbolic feedback
  \item Triadic knowledge graph structures
\end{itemize}

\subsection*{Metrics}

\begin{itemize}
  \item Mutual information estimation (MINE, CLUB)
  \item BERTScore and BLEU for semantic coherence
  \item Cosine similarity of embeddings over recursion steps
  \item Entropy tracking over recursive depth
\end{itemize}

\section*{V. Visualization}

\begin{itemize}
  \item Trajectory plots of $\Phi_r(t)$ and $\Delta I(t)$
  \item UMAP or t-SNE projections of embedding drift
  \item Recursion density graphs from symbolic logic cycles
\end{itemize}

\section*{VI. Cognitive Analogy}

\begin{itemize}
  \item Circular reasoning in argument or belief systems
  \item Mental fatigue from recursive contradiction processing
  \item Symbolic collapse under semantic overload
\end{itemize}

\section*{VII. Applications}

\begin{itemize}
  \item AI architecture design with recursion safety constraints
  \item Symbolic knowledge graph collapse prevention
  \item Cognitive model simulations of contradiction handling
  \item Collapse detection in triadic symbolic systems
\end{itemize}

\section*{VIII. Future Work}

\begin{itemize}
  \item Extend theorem to $n$-node mutual implication loops
  \item Develop symbolic dampening mechanisms for recursive feedback
  \item Integrate detection into live graph-based AI systems
\end{itemize}

\end{document}